---
title: "Reproducible Research, Course Project 1"
author: "Maya Ivanova"
date: "May 7, 2017"
output:
  md_document: default
  html_document: default
  pdf_document: default
---
Input data
```{r}
activity <- read.csv("activity.csv", header = TRUE, sep = ",")
head(activity)
```

 
```{r}
#Calculate the total number of steps taken each day.
#I will use aggregate function to calculate sum, mean, and median number of steps
#I will Ignore missing values with na.rm = T
StepsByDay <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
```
Histogram of the total number of steps taken each day
```{r}
require("RColorBrewer")
hist(StepsByDay$steps, main = "Steps by Day", xlab = "day", col = topo.colors(6)) 
```

Calculate the mean and median number of steps taken each day
```{r}
aggregate(steps ~ date, data = activity, FUN = mean, na.rm =TRUE)
aggregate(steps ~ date, data = activity, FUN = median, na.rm =TRUE)
```
Time series plot of the average number of steps taken
```{r}
#Make time series plot of the 5-minute interval (x-axis) and the average number
#of steps taken, averaged across all days (y-axis)

time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(row.names(time_series), time_series, type = "l", xlab = "Interval", 
     ylab = "Average number of steps", main = "Average number of steps over all days",
     col = "blue")
```
The 5-minute interval that, on average, contains the maximum number of steps
```{r}
max_interval <- which.max(time_series)
names(max_interval)
```
Imputing missing data
```{r}
#Note that there are a number of days/intervals where there are missing values
#(coded as NA). The presence of missing days may introduce bias into some 
#calculations or summaries of the data. We will calculate and report the total
#number of missing values in the dataset #(i.e. the total number of rows with NAs)

activity_NA <- sum(is.na(activity))
activity_NA

#Devise a strategy for filling in all of the missing values in the dataset.
StepsAverage <- aggregate(steps ~ interval, data = activity, FUN = mean)
fill_NA <- numeric()
for (i in 1:nrow(activity)) {
        na <- activity[i, ]
        if (is.na(na$steps)) {
                steps <- subset(StepsAverage, interval == na$interval)$steps
        } else {
                steps <- na$steps
        }
        fill_NA <- c(fill_NA, steps)
}

#Create a new dataset that is equal to the original dataset but with 
#the missing data filled in.
new_activity <- activity
new_activity$steps <- fill_NA
```
Make a histogram of the total number of steps taken each day after missing values are imputed
```{r}
new_StepsByDay <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)

#Create a histogram
require("RColorBrewer")
hist(new_StepsByDay$steps, main = "Total Steps by Day", xlab = "day", col = topo.colors(6))
```
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
#Create a new  a new variable "type_of_day" indicating whether a 
#given date is a weekday or #weekend day.
new_activity['type_of_day'] <- weekdays(as.Date(new_activity$date))
new_activity$type_of_day[new_activity$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
new_activity$type_of_day[new_activity$type_of_day != "weekend"] <- "weekday"

# convert "type_of_day" from character to factor
new_activity$type_of_day <- as.factor(new_activity$type_of_day)

# calculate average steps by interval across all days

df_imputed_steps_by_interval <- aggregate(steps ~ interval + type_of_day, new_activity, mean)

#create a plot
require(ggplot2)
plot<- ggplot(df_imputed_steps_by_interval, aes(x =interval , y=steps, color=type_of_day)) +
        geom_line() +
        labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") +
        facet_wrap(~ type_of_day, ncol = 1, nrow=2)
print(plot)
```



